About p-adic stability

### 2014-12-10 examples/stability.tex
Smith form for matrices

Differentiating ring operations (sum, products, inverse) over matrices 
is again straightforward, keeping in mind that matrix algebras are not
commutative.

\subsection*{Determinants and characteristic polynomials.}

We first outline the standard computation of the differential of the function
$\det : M_n(K) \to K$. Suppose that $M \in \GL_n(K)$ and that $\com(M) = \det(M) M^{-1}$.
Then
\begin{align*}
\det(M + dM) &= \det(M) \cdot \det(I + M^{-1} \cdot dM)  \\
&= \det(M) \cdot \big(1 + \tr(M^{-1} \cdot dM)\big) \\
&= \det(M) + \tr(\com(M) \cdot dM).
\end{align*}
The differential of $\det$ at $M$ is then $dM \mapsto \tr(\com(M) \cdot dM)$. It turns
out that this formula is still valid when $M$ is not invertible.
The same computation extends readily to characteristic polynomials,
since they are defined as determinants. More precisely, let us 
consider the function $\chi : M_n(K) \to K_n[X]$ taking a matrix 
$M$ to its monic characteristic polynomial $\det(X-M)$.
Then $\chi$ is differentiable at each point $M \in M_n(K)$ and its 
differential is given by $dM \mapsto \tr(\com(X{-}M) \cdot dM)$.

@@

\subsection{Stability}

>>>>

\todo{Write a short introduction.}

The notation $M_{n,m}(K)$ refers to the $K$-vector space of matrices 
with $n$ rows and $m$ columns with coefficients in $K$.

\noindent
{\bf Smith form.}
In what follows, we will often use the \emph{Smith decomposition} of a 
matrix $M \in M_{n,m}(K)$: we recall that it is a factorization of $M$ 
as a product $U_M \cdot \Delta_M \cdot V_M$ when $U_M$ and $V_M$ are 
unimodular (square) matrices of size $n$ and $m$ respectively, with 
coefficients in $\O_K$ and $\Delta_M$ satisfies the two following
conditions:

\noindent
(i) its $(i,j)$-th entry vanishes as soon as $i \neq j$

\noindent
(ii) its $(i,i)$-th entry divides its $(j,j)$-th entry when $i
\leq j$.

\noindent
Such a factorization is unique and can be easily computed by performing 
elementary operations on rows and columns.
Regarding to precision, one can prove that if the entries of $M$ are all 
known up to some $O(\pi^N)$, it is possible to compute a factorization 
$U_M \cdot \Delta_M \cdot V_M$ where $U_M$ and $V_M$ are \emph{exact}
matrices satisfying the above conditions and where $\Delta_M$ is known
at precision $O(\pi^N)$ and satisfies~(i) and~(ii) modulo $\pi^N$.

\subsection{Product of matrices}

\subsection{Determinant}

We first outline the standard computation of the differential of the function
$\det : M_n(K) \to K$. Suppose that $M \in \GL_n(K)$ and that $\com(M) = \det(M) M^{-1}$.
Then
\begin{align*}
\det(M + dM) &= \det(M) \cdot \det(I + M^{-1} \cdot dM)  \\
&= \det(M) \cdot \big(1 + \tr(M^{-1} \cdot dM)\big) \\
&= \det(M) + \tr(\com(M) \cdot dM).
\end{align*}
The differential of $\det$ at $M$ is then $dM \mapsto \tr(\com(M) \cdot dM)$. It turns
out that this formula is still valid when $M$ is not invertible.

\subsection{Characteristic polynomials.}

The same computation extends readily to characteristic polynomials,
since they are defined as determinants. More precisely, let us 
consider the function $\chi : M_n(K) \to K_n[X]$ taking a matrix 
$M$ to its monic characteristic polynomial $\det(X-M)$.
Then $\chi$ is differentiable at each point $M \in M_n(K)$ and its 
differential is given by $dM \mapsto \tr(\com(X{-}M) \cdot dM)$.

@@

\subsection{Application to precision}
